---
title: "ML-prjoect"
author: "MD"
date: "September 23, 2018"
output: html_document
---

## Overview
####This document encompasses Coursera's course Practical Machine Learning project report. This project is aimed predicting the manner in which 6 participants performed some exercise . Using specific devices (Jawbone up, Nike FuelBand...) which are part of quantified self movement can collect data ata about personal activity. More details on gathering data for this study is located a : http://groupware.les.inf.puc-rio.br/har 



##A-Uploading data and exploratory Analysis
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
#required library
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(12345)

# Training data source link :https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
# Testing data source link :https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

#Read data sets from URLs

MLtraining <- read.csv((url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")), na.strings= c("NA", ""))
MLtesting <- read.csv((url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")), na.strings= c("NA", ""))

# create a partition with the training dataset 
inTrain  <- createDataPartition(MLtraining$classe, p=0.7, list=FALSE)
TrainSet <- MLtraining[inTrain, ]
TestSet  <- MLtraining[-inTrain, ]

#data sets clean up : Variables with NA have to be removed 
#NZV (Near Zero variance) and ID variables will be removed.

# NZV variables clean up 
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]

# NA variables removal
dataNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, dataNA==FALSE]
TestSet  <- TestSet[, dataNA==FALSE]

# Identification variables (columns 1 to 5) clean up
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]

#Depict matrix correlation 
correlatMatrix <- cor(TrainSet[, -54])
corrplot(correlatMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))


```


```{r SumStepsperday}
##B-Prediction Models
##Prediction models will be as follow:a-Random Forest; b-Decision Tree; c-Generalized Boosted Model

#a) Method: Random Forest
# model fit
set.seed(12345)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel

# prediction on Test dataset
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest

#b) Method: Decision Trees
# model fit
set.seed(12345)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modFitDecTree)

# prediction on Test dataset
predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree

##c) Method: Generalized Boosted Model
# model fit
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel

# prediction on Test dataset
predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM


```

##C. Applying the Selected Model to the Test Data
####The accuracy of the 3 regression modeling methods above are:
####a.	Random Forest : 0.9992
####b.	Decision Tree : 0.7208
####c.	GBM : 0.9856
####Therefore, Random Forest model will be applied to predict (testing dataset) as follow.

```{r Daily Total steps, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
predictTEST <- predict(modFitRandForest, newdata=MLtesting)
predictTEST

```
